---
title: "Data Exploration"
---

```{r setup.1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "images/")
```

```{r setup.2}
library(data.table, warn.conflicts = FALSE)
library(igraph, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
```

## Whole data set

```{r test}
total_info <- read.delim("data/raw/arg_all/arg_datasets.tsv", sep = "\t", header = TRUE, fill = TRUE )
colnames(total_info)[c(9, 10, 12)] <- c("environment_package", "seq_type", "mean_seq_length")
head(total_info)
```


```{r biome_info.1}
biome_info_2 <- table(unlist(total_info$biome))
ggplot(as.data.frame(biome_info_2[biome_info_2 > 100]), aes(x = reorder(Var1, Freq), y = Freq)) + geom_bar(stat = "identity") +
    labs(x = element_blank(), y = element_blank(), title = "Biome counts: All datasets (n = 7848)") +
    theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r biome_info.2}
biome_info <- read.table("data/raw/arg_all/arg_datasets.biomes.tsv", sep = "\t", header = TRUE,
                         col.names = c("counts", "biome"))

ggplot(data.frame(biome_info), aes(x = biome, y = counts)) + geom_bar(stat = "identity") +  
  labs(x = element_blank(), y = element_blank(), title = "Biome counts: All datasets (n = 7848)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r env_info.1}
env_info <- read.table("data/raw/arg_all/arg_datasets.envs.tsv", sep = "\t", header = TRUE,
                         col.names = c("counts", "environment"))

ggplot(data.frame(env_info), aes(x = environment, y = counts)) + geom_bar(stat = "identity") +  
  labs(x = element_blank(), y = element_blank(), title = "Environment counts: All datasets (n = 7848)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, hjust = 1))
```
```{r env_info.2}
env_info_2 <- table(unlist(total_info$environment_package))
ggplot(as.data.frame(env_info_2[env_info_2 > 50]), aes(x = reorder(Var1, Freq), y = Freq)) + geom_bar(stat = "identity") +
    labs(x = element_blank(), y = element_blank(), title = "Environment counts: All datasets (n = 7848)") +
    theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, hjust = 1))

```

## Reads Info

```{r reads.1}

```

## Subset

```{r read_subset_info.1}
arg_subset_info <- read.table("data/raw/arg_subsets/arg_datasets.subset.tsv", sep = "\t", header = TRUE)
head(arg_subset_info)
```

```{r add_arg_ratio_column.3}
proportion = 0.1
cutoff = quantile((sort(arg_subset_info$arg_percent)), 1 - proportion)
```

```{r add_arg_ratio_column.2}
ggplot(data.frame(val = arg_subset_info$arg_percent), aes(val)) + geom_histogram(binwidth = 0.001) +
  labs(x = "ARG Percentage", y = "Count", title = "ARG proportion counts (n = 934)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_vline(xintercept = 4.545e-04, colour = "#a9f9cd", linetype = "dashed") +
  geom_vline(xintercept = cutoff, colour = "#BB0000", linetype = "dashed") 
```
## Cooccurrence Data

`arg_cooccurrence.tsv` contains the cooccurrence data (ie, [Spearman' correlation](https://en.wikipedia.org/wiki/Spearman%27s_rank_correlation_coefficient)) for all the ARGs found in `arg_datasets.subset.tsv`.

```{r load_cooccurrence_data.1}
arg_cooccurrence_table <- read.table("data/raw/cooccurrences/arg_cooccurrence.tsv", 
                                col.names = c("ARG_1", "ARG_2", "rho", "p"))
head(arg_cooccurrence_table)
```

Since all the cooccurrence values we have here are significant - that is, with $p < 0.05$ - we will remove the $p$ value column from `arg-cooccurrence_table`:

```{r load_cooccurrence_data.2}
arg_cooccurrence_table <- arg_cooccurrence_table[, -4]
head(arg_cooccurrence_table)
```

## Graph Analysis

We begin our network analysis of the ARG cooccurrence by separating `arg_cooccurrence_table` by positive and negative values and then:

1. Finding the optimal community structure of each observed network 
2. Generating mock community structure of each network and recording the modularity of each mock community structure, as well as which ARGs are the most central
3. Comparing the modularity of the observed network against the distribution of the mock communities to determine statistical significance

```{r graph_analysis.1}
positive_arg_cooccurrence <- arg_cooccurrence_table[arg_cooccurrence_table[['rho']] > 0, ]
gr_pos_cooccurrence <- graph.data.frame(positive_arg_cooccurrence, directed = FALSE)
negative_arg_cooccurrence <- arg_cooccurrence_table[arg_cooccurrence_table[['rho']] < 0, ]
gr_neg_cooccurrence <- graph.data.frame(positive_arg_cooccurrence, directed = FALSE)
```

We'll now 

```{r graph_analysis.2}
plot(gr_pos_cooccurrence, vertex.size = 3, vertex.label = NA)
```

Finding communities.  Beginning with greedy optimization to get big picture.

```{r graph_analysis.3}
show_community_info <- function(graph_object){
  communities_object <- cluster_fast_greedy(graph_object, weights = E(graph_object)$rho)
  hot_otus <- vector
  for (community in communities(communities_object)){
    sub_graph <- induced.subgraph(graph = graph_object, 
                                    v = community, impl = "auto")
    highest_degree_otu <- V(sub_graph)[which.max(degree(sub_graph))]$name
    hot_otus <- c(hot_otus, highest_degree_otu)
    print(paste("Highest degree OTU was", highest_degree_otu, "(degree =",  
              degree(sub_graph, highest_degree_otu), "):"))
  }
  plot(communities_object, graph_object, 
     layout = layout.lgl,
     rescaled = TRUE,
     vertex.size = ifelse(V(graph_object) %in% hot_otus, 5, 1), 
     vertex.label = ifelse(V(graph_object) %in% hot_otus, V(graph_object)$id, NA),
     vertex.label.cex = 2
     )
  modularity(graph_object, membership(communities_object))
}
```

### Low ARG percentage

```{r low.cooc.1}
low.arg_subset_info <- read.table("data/raw/arg_subsets/arg_datasets.subset.low.tsv", sep = "\t", header = TRUE)
ggplot(data.frame(low.arg_subset_info), aes(biome)) + geom_bar() +
  labs(x = element_blank(), y = element_blank(), title = "Biome counts: Low ARG Prevalence (n = 700)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r low.cooc.2}
ggplot(data.frame(low.arg_subset_info), aes(environment_package)) + geom_bar() +
  labs(x = element_blank(), y = element_blank(), title = "Environment counts: Low ARG Prevalence (n = 700)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r low.1}
arg_cooccurrence.low <- read.table("data/raw/cooccurrences/arg_cooccurrence.low.tsv", 
                                col.names = c("ARG_1", "ARG_2", "rho", "p"))
arg_cooccurrence.low <- arg_cooccurrence.low[, -4]
head(arg_cooccurrence.low)
```

```{r low.2}
low.pos_cooccurrence <- arg_cooccurrence.low[arg_cooccurrence.low[['rho']] > 0, ]
low.positive_graph <- graph.data.frame(low.pos_cooccurrence, directed = FALSE)
low.neg_cooccurrence <- arg_cooccurrence.low[arg_cooccurrence.low[['rho']] <= 0, ]
low.neg_cooccurrence[[3]] <- abs(low.neg_cooccurrence[[3]])
low.negative_graph <- graph.data.frame(low.neg_cooccurrence, directed = FALSE)
```

```{r low.3}
low.mod_pos <- show_community_info(low.positive_graph)
```

```{r low.4}
low.mod_neg <- show_community_info(low.negative_graph)
```

```{r low.5}
vertex_count <- vcount(low.positive_graph)
edge_count <- ecount(low.positive_graph)
low.bootstraps <- vector()

for (i in 1:1000)
{
  g <- erdos.renyi.game(vertex_count, edge_count, type = "gnm")
  E(g)$rho <- runif(length(E(g)), 0, 1)
  g_communities <- cluster_fast_greedy(g, weights = E(g)$rho)
  low.bootstraps <- c(low.bootstraps, modularity(g, membership(g_communities)))
}

low.bootstrap_cutoff <- quantile((sort(low.bootstraps)), 0.95)

ggplot(data.frame(val = low.bootstraps), aes(val)) + geom_histogram(binwidth = 0.0005) +
  labs(x = "Modularity", y = "Counts", title = "Modularity scores of simulated networks (n = 1000)",
       subtitle = "(Low ARG environments)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
  geom_vline(xintercept = low.bootstrap_cutoff, colour = "#BB0000", linetype = "dashed")
```

```{r low.6}
ggplot(data.frame(val = low.bootstraps), aes(val)) + geom_histogram(binwidth = 0.0005) +
  labs(x = "Modularity", y = "Counts", title = "Modularity scores of simulated networks (n = 1000)",
       subtitle = "(Low ARG environments)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
  geom_vline(xintercept = low.bootstrap_cutoff, colour = "#BB0000", linetype = "dashed") +
  geom_vline(xintercept = low.mod_pos, colour="#ed64f4", linetype="dashed") +
  geom_vline(xintercept = low.mod_neg, colour = "#f4ad42", linetype = "dashed") 
```


### Medium ARG percentage

```{r med.cooc.1}
med.arg_subset_info <- read.table("data/raw/arg_subsets/arg_datasets.subset.med.tsv", sep = "\t", header = TRUE)
ggplot(data.frame(med.arg_subset_info), aes(biome)) + geom_bar() +
  labs(x = element_blank(), y = element_blank(), title = "Biome counts: Medium ARG Prevalence (n = 140)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, hjust = 1))
```
```{r med.cooc.2}
ggplot(data.frame(med.arg_subset_info), aes(environment_package)) + geom_bar() +
  labs(x = element_blank(), y = element_blank(), title = "Environment counts: Medium ARG Prevalence (n = 140)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r med.1}

arg_cooccurrence.med <- read.table("data/raw/cooccurrences/arg_cooccurrence.med.tsv", 
                                col.names = c("ARG_1", "ARG_2", "rho", "p"))
arg_cooccurrence.med <- arg_cooccurrence.med[, -4]
head(arg_cooccurrence.med)
```

```{r med.2}
med.pos_cooccurrence <- arg_cooccurrence.med[arg_cooccurrence.med[['rho']] > 0, ]
med.positive_graph <- graph.data.frame(med.pos_cooccurrence, directed = FALSE)
med.neg_cooccurrence <- arg_cooccurrence.med[arg_cooccurrence.med[['rho']] <= 0, ]
med.neg_cooccurrence[[3]] <- abs(med.neg_cooccurrence[[3]])
med.negative_graph <- graph.data.frame(med.neg_cooccurrence, directed = FALSE)
```

```{r med.3}
med.mod_pos <- show_community_info(med.positive_graph)
```

```{r med.4}
med.mod_neg <- show_community_info(med.negative_graph)
```


```{r med.5}
vertex_count <- vcount(med.positive_graph)
edge_count <- ecount(med.positive_graph)
med.bootstraps <- vector()

for (i in 1:1000)
{
  g <- erdos.renyi.game(vertex_count, edge_count, type = "gnm")
  E(g)$rho <- runif(length(E(g)), 0, 1)
  g_communities <- cluster_fast_greedy(g, weights = E(g)$rho)
  med.bootstraps <- c(med.bootstraps, modularity(g, membership(g_communities)))
}

med.bootstrap_cutoff <- quantile((sort(med.bootstraps)), 0.95)

ggplot(data.frame(val = med.bootstraps), aes(val)) + geom_histogram(binwidth = 0.0005) +
  labs(x = "Modularity", y = "Counts", title = "Modularity scores of simulated networks (n = 1000)",
       subtitle = "(Medium ARG environments)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
  geom_vline(xintercept = med.bootstrap_cutoff, colour = "#BB0000", linetype = "dashed")
```

```{r med.6}
ggplot(data.frame(val = med.bootstraps), aes(val)) + geom_histogram(binwidth = 0.0005) +
  labs(x = "Modularity", y = "Counts", title = "Modularity scores of simulated networks (n = 1000)",
       subtitle = "(Medium ARG environments)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
  geom_vline(xintercept = med.bootstrap_cutoff, colour = "#BB0000", linetype = "dashed") +
  geom_vline(xintercept = med.mod_pos, colour="#ed64f4", linetype="dashed") +
  geom_vline(xintercept = med.mod_neg, colour = "#f4ad42", linetype = "dashed") 
```

### High ARG percentage

```{r high.cooc.1}
high.arg_subset_info <- read.table("data/raw/arg_subsets/arg_datasets.subset.high.tsv", sep = "\t", header = TRUE)
ggplot(data.frame(high.arg_subset_info), aes(biome)) + geom_bar() +
  labs(x = element_blank(), y = element_blank(), title = "Biome counts: High ARG Prevalence (n = 94)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 135, hjust = 1))
```

```{r high_sets}
high.arg_subset_info[order(high.arg_subset_info$arg_percent, decreasing = TRUE), ]
```

```{r high.cooc.2}
ggplot(data.frame(high.arg_subset_info), aes(environment_package)) + geom_bar() +
  labs(x = element_blank(), y = element_blank(), title = "Biome counts: High ARG Prevalence (n = 94)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 135, hjust = 1))
```


```{r high.1}

arg_cooccurrence.high <- read.table("data/raw/cooccurrences/arg_cooccurrence.high.tsv", 
                                col.names = c("ARG_1", "ARG_2", "rho", "p"))
arg_cooccurrence.high <- arg_cooccurrence.high[, -4]
head(arg_cooccurrence.high)
```

```{r high.2}
high.pos_cooccurrence <- arg_cooccurrence.high[arg_cooccurrence.high[['rho']] > 0, ]
high.positive_graph <- graph.data.frame(high.pos_cooccurrence, directed = FALSE)
high.neg_cooccurrence <- arg_cooccurrence.high[arg_cooccurrence.high[['rho']] <= 0, ]
high.neg_cooccurrence[[3]] <- abs(high.neg_cooccurrence[[3]])
high.negative_graph <- graph.data.frame(high.neg_cooccurrence, directed = FALSE)
```

```{r high.3}
high.mod_pos <- show_community_info(high.positive_graph)
```

The mexF gene is highly important, which corresponds with findings that it enhances fitness.

```{r high.4}
high.mod_neg <- show_community_info(high.negative_graph)
```

```{r high.5}
vertex_count <- vcount(high.positive_graph)
edge_count <- ecount(high.positive_graph)
high.bootstraps <- vector()

for (i in 1:1000)
{
  g <- erdos.renyi.game(vertex_count, edge_count, type = "gnm")
  E(g)$rho <- runif(length(E(g)), 0, 1)
  g_communities <- cluster_fast_greedy(g, weights = E(g)$rho)
  high.bootstraps <- c(high.bootstraps, modularity(g, membership(g_communities)))
}

high.bootstrap_cutoff <- quantile((sort(high.bootstraps)), 0.95)

ggplot(data.frame(val = high.bootstraps), aes(val)) + geom_histogram(binwidth = 0.0005) +
  labs(x = "Modularity", y = "Counts", title = "Modularity scores of simulated networks (n = 1000)",
       subtitle = "(High ARG environments)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
  geom_vline(xintercept = high.bootstrap_cutoff, colour = "#BB0000", linetype = "dashed")
```

```{r high.6}
ggplot(data.frame(val = high.bootstraps), aes(val)) + geom_histogram(binwidth = 0.0005) +
  labs(x = "Modularity", y = "Counts", title = "Modularity scores of simulated networks (n = 1000)",
       subtitle = "(High ARG environments)") +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
  geom_vline(xintercept = high.bootstrap_cutoff, colour = "#BB0000", linetype = "dashed") +
  geom_vline(xintercept = high.mod_pos, colour="#ed64f4", linetype="dashed") +
  geom_vline(xintercept = high.mod_neg, colour = "#f4ad42", linetype = "dashed") 
```

