---
title: "Data Exploration"
---

```{r setup.1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup.2}
library(data.table, warn.conflicts = FALSE)
library(igraph, warn.conflicts = FALSE)
```

## Whole data set

```{r load_complete_metag_data.1}

```

## Subset

## Cooccurrence Data

`arg_cooccurrence.tsv` contains the cooccurrence data (ie, [Spearman' correlation](https://en.wikipedia.org/wiki/Spearman%27s_rank_correlation_coefficient)) for all the ARGs found in `arg_datasets.subset.tsv`.

```{r load_cooccurrence_data.1}
arg_cooccurrence_table <- read.table("data/raw/arg_cooccurrence.tsv", 
                                col.names = c("ARG_1", "ARG_2", "rho", "p"))
head(arg_cooccurrence_table)
```
Since all the cooccurrence values we have here are significant - that is, with $p < 0.05$ - we will remove the $p$ value column from `arg-cooccurrence_table`:

```{r load_cooccurrence_data.2}
arg_cooccurrence_table <- arg_cooccurrence_table[, -4]
head(arg_cooccurrence_table)
```

## Graph Analysis

We begin our network analysis of the ARG cooccurrence by:
1. Separating `arg_cooccurrence_table` into positive and negative values, t
2. 

```{r graph_analysis.1}
positive_arg_cooccurrence <- arg_cooccurrence_table[arg_cooccurrence_table[['rho']] > 0, ]
negative_arg_cooccurrence <- arg_cooccurrence_table[arg_cooccurrence_table[['rho']] < 0, ]
```

We'll now 

```{r graph_analysis.2}
gr_pos_cooccurrence <- graph.data.frame(positive_arg_cooccurrence, directed = FALSE)
plot(gr_pos_cooccurrence, vertex.size = 3, vertex.label = NA)

```

Finding communities.  Beginning with greedy optimization to get big picture.

```{r graph_analysis.3}
  
  communities_object <- cluster_fast_greedy(gr_pos_cooccurrence, weights = E(gr_pos_cooccurrence)$rho)
  hot_otus <- vector
  for (community in communities(communities_object)){
    sub_graph <- induced.subgraph(graph = gr_pos_cooccurrence, 
                                    v = community, impl = "auto")
    highest_degree_otu <- V(sub_graph)[which.max(degree(sub_graph))]$name
    hot_otus <- c(hot_otus, highest_degree_otu)
    print(paste("Highest degree OTU was", highest_degree_otu, "(degree =",  
              degree(sub_graph, highest_degree_otu), "):"))
  }
  plot(communities_object, gr_pos_cooccurrence, 
     layout = layout.fruchterman.reingold(gr_pos_cooccurrence),
     rescaled = TRUE,
     vertex.size = ifelse(V(gr_pos_cooccurrence) %in% hot_otus, 5, 1), 
     vertex.label = ifelse(V(gr_pos_cooccurrence) %in% hot_otus, V(gr_pos_cooccurrence)$id, NA),
     vertex.label.cex = 2
     )
```

```{r graph_analysis.4}
modularity(gr_pos_cooccurrence, membership(communities_object))
```

