---
title: "tda.Rmd"
---

```{r setup.1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup.2, warning = FALSE }
library(ecodist, warn.conflicts = FALSE)
library(ggplot2, warn.conflicts = FALSE)
library(MASS, warn.conflicts = FALSE)
library(TDA, warn.conflicts = FALSE)
library(TDAstats, warn.conflicts = FALSE)
library(phyloseq, warn.conflicts = FALSE)
library(phylosmith, warn.conflicts = FALSE)
library(RColorBrewer, warn.conflicts = FALSE)
library(tidyverse, warn.conflicts = FALSE)
library(TDA, warn.conflicts = FALSE)
library("TDAstats", warn.conflicts = FALSE)
library(vegan, warn.conflicts = FALSE)
```

# Persistent homology

We'll begin our exploration of topological data analysis with the application of persistent homology.  We'll start on the soil column data set, as there has been a clear signal demonstrated in NMDS plots indicating separation between soil-treated and manure-treated columns.  Persistent homology should be able to detect some difference in the topological structure of two different treatments. After this verification, we will move on to applying persistent homology to our data set.

## Soil column 

We'll begin by reading in the soil column data sets.  

```{r read.arg}
soil_column.args <- readRDS("./data/raw/arg-phyloseq.RDS")
soil_column.args
```

```{r get.cooccur}
soil_column.args.cooccur <- FastCoOccur(soil_column.args, "Treatment")

```

```{r}
phyloseq_ggplot_NMDS <- function(phyloseq_obj, treatment){
  require(ggplot2); require(RColorBrewer); require(vegan); require(MASS); require(phylosmith)
  getPalette = colorRampPalette(brewer.pal(8, "Dark2")); colorCount = 1 + length(unlist(unique(sample_data(phyloseq_obj)[[treatment]]))); colors = getPalette(colorCount); theme_set(theme_bw())

  MDS <- metaMDS(t(otu_table(find_generalists(phyloseq_obj))), autotransform = FALSE, distance = "bray", k=3, trymax=50)
  plot(MDS, display = c("sites", "species"), choices = c(1,2), type = "p")
  abline(h=0,lty=2)
  abline(v=0,lty=2)
  stressplot(MDS)

  Treatment <- sample_data(phyloseq_obj)[[treatment]]
  MDS$points[,'MDS1']
  MDS1 <- data.frame(scores(MDS))$NMDS1
  MDS2 <- data.frame(scores(MDS))$NMDS2
  NMDS <- data.frame(MDS1,MDS2,Treatment)
  NMDS.narm <- subset(NMDS, !is.na(Treatment))
  veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100){
    theta <- (0:npoints) * 2 * pi/npoints
    Circle <- cbind(cos(theta), sin(theta))
    t(center + scale * t(Circle %*% chol(cov)))}
  df_ell <- data.frame()
  for(g in unique(NMDS.narm[[treatment]])){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS.narm[NMDS.narm[,treatment]==g,], veganCovEllipse(cov.wt(cbind(MDS1,MDS2),wt=rep(1/length(MDS1),length(MDS1)))$cov,center=c(mean(MDS1),mean(MDS2))))),group=g))}

  p <- ggplot(data = NMDS.narm, aes(MDS1, MDS2)) +
    geom_point(aes(color = treatment), size=1.5, alpha=0.75) +
    geom_path(data=df_ell, aes(x=MDS1, y=MDS2, colour=group), size=2, linetype=5) +
    scale_color_manual(values=colors) +
    theme_classic() +
    theme(aspect.ratio=1,
          axis.line.x = element_line(colour = 'black', size = 1, linetype = 'solid'),
          axis.line.y = element_line(colour = 'black', size = 1, linetype = 'solid'),
          axis.text.x = element_text(size = 10, face = "bold"),
          axis.text.y = element_text(size = 10, face = "bold"),
          axis.title.x = element_text(size = 12, face = "bold"),
          axis.title.y = element_text(size = 12, face = "bold"),
          legend.title = element_blank(),
          legend.text = element_text(size = 11, face = "bold"),
          legend.background = element_rect(fill = (alpha = 0))
    )
  return(p)
}
phyloseq_ggplot_NMDS(soil_column.args, "Treatment")
```

Performing NMDS, reducing data.  Preparing data for persistent homology.

```{r tda.1}
soil_column.nmds <- metaMDS(t(otu_table(find_generalists(soil_column.args))), autotransform = FALSE,
                            distance = "bray", k = 2, trymax = 50)
PCA_1 <- data.frame(scores(soil_column.nmds))$NMDS1
PCA_2 <- data.frame(scores(soil_column.nmds))$NMDS2
treatments <- sample_data(soil_column.args)[["Treatment"]]
soil_column.nmds.transform <- data.frame(PCA_1, PCA_2, treatments)
soil_column.nmds.transform.narm <- subset(soil_column.nmds.transform, !is.na(treatments))
```

Separating data by category.

```{r tda.2}
soil_column.nmds.manure <- soil_column.nmds.transform.narm %>% 
  filter(treatments == "Manure") %>% 
  select(PCA_1, PCA_2) %>% 
  as.matrix()
soil_column.nmds.control <- soil_column.nmds.transform.narm %>% 
  filter(treatments == "Control") %>% 
  select(PCA_1, PCA_2) %>% 
  as.matrix()
```

Persistent homology by treatment.

```{r tda.3}
manure.hom <- calculate_homology(soil_column.nmds.manure, dim = 3)
plot_persist(manure.hom)
```
```{r}
plot_barcode(manure.hom)
```
```{r}
plot(manure.hom)
```

```{r}
ggplot(data = soil_column.nmds.transform.narm, aes(x = PCA_1, y = PCA_2, color = treatments)) +
  geom_point()
```


